---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

To load the data use this code:

```{r loading and preprocesing, echo=TRUE}

data<-read.csv(unz("activity.zip",filename = "activity.csv"))

```

## What is mean total number of steps taken per day?

1. The total number of steps taken per day is
```{r mean total number, echo=TRUE}
ts<-tapply(data$steps,data$date,sum,na.rm=TRUE)

```

2. the  histogram of the total number of steps taken each day is:

```{r histogram, echo=TRUE}

hist(ts)

```

3. the mean and median are:

```{r mean and median,echo=TRUE, results="asis"}

mean_ts<-mean(ts)
median_ts<-median(ts)
df<-data.frame(mean_ts,median_ts)
names(df)<-c("mean","median")

library(xtable)
print(xtable(df),type = "html")

```


## What is the average daily activity pattern?

1- the time series plot for the average number of steps taken averaged across all days is:

```{r time serie,echo=TRUE}

tsa<-tapply(data$steps,data$interval,mean,na.rm=TRUE)
interval<-data$interval[1:288]

plot(interval,tsa,type = "l")

```

2. the 5-minute interval whit the maximum number of sterp is:

```{r maximun in 5 minute interval, echo=TRUE}

interval[which.max(tsa)]

```


## Imputing missing values

the total number of missing values in the dataset is:

```{r missing values,echo=TRUE}

sum(is.na(data$steps))

```

to replace the missing values we are going to use the 5-minute intervale average, thus:

```{r replace missing values,echo=TRUE}

library(dplyr)
data2<-data

impute<-function(x,fun){
        missing<-is.na(x)
        replace(x,missing,fun(x[!missing]))
}

data2<-data %>% group_by(interval) %>% mutate(steps=impute(steps,mean))
```



The new total number of steps taken per day is
 
```{r mean total number2, echo=TRUE}
ts2<-tapply(data2$steps,data2$date,sum,na.rm=TRUE)

```

the new histogram of the total number of steps taken each day is:

```{r histogram2, echo=TRUE}

hist(ts2)

```

the new mean and median are:

```{r mean and median 2,echo=TRUE, results="asis"}

mean_ts2<-mean(ts2)
median_ts2<-median(ts2)
df2<-data.frame(mean_ts2,median_ts2)
names(df2)<-c("mean","median")

library(xtable)
print(xtable(df2),type = "html")
```


## Are there differences in activity patterns between weekdays and weekends?

Creating factor variable weekday or weekend

```{r creating factor,echo=TRUE}

data2$date <- as.Date(data2$date)
weekdays1 <- c("lunes", "martes", "miércoles", "jueves", "viernes")

data2$wDay <- factor((weekdays(data2$date) %in% weekdays1), levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))

```


```{r plot,echo=TRUE}

library(lattice)

data3<-data2 %>% group_by(wDay,interval) %>% summarize(mean=mean(steps))

xyplot(mean~interval | wDay,data = data3,layout=c(1,2), type="l")

```